### 2011-11-07 doc/pldi2012/circuit.tex
circuit

The circuit simulation takes a collection of
of wires and nodes where wires meet.  
The simulation repeatedly updates
currents, distributes charges, and updates voltages for as many
time steps as the simulation demands.  We first
describe the partitioning of regions in Section~\ref{subsec:partitioning}
and then describe the data type declarations 
and the constraints they enforce on the simulation that enable
the runtime to efficiently execute the application
in Section~\ref{subsec:datatypes}.

@@

Lines 19-27 are responsible for partitioning the circuit into {\tt MAX\_PIECES}
pieces that can be worked on in parallel.  To partition a region, we
provide a {\em coloring}, which is a relation on the elements
of a region and a set of colors.  A {\em partition} is an object which given
a coloring and a region $r$ contains for each color $c$ a {\em subregion} of $r$
with all the elements of $r$ colored $c$.

The region of all nodes in the circuit is first partitioned into 
{\em private} and {\em shared} nodes (lines 20-21).  Private nodes
are touched by a single task in a phase, while shared nodes may
be referenced by multiple tasks in a phase.  The thick lines in figure \ref{sfig:part_fig:pvs}
show how the nodes in a small graph might be partitioned into three subsets.  Gray
nodes are shared and black nodes are private.  Note that
this partition is disjoint since each node has one color.
%Our current implementation uses small
%integers or booleans for the names of colors, but any enumerable set would be
%fine.  The coloring relation need not be total - it can leave out some elements entirely.  And although it is not required to be injective (it can map a single
%element to multiple colors), the fact that injective mappings will always
%result in disjoint subregions makes them strongly encouraged (when possible).
%Our example uses four colorings to create circuit partitions.

After partitioning into all the nodes that will be private and shared, we now
need to partition each of these subregions to generate {\tt MAX\_PIECES}-way
parallelism.  We first parition the private nodes into each of the disjoint
{\tt MAX\_PIECES} subregions (lines 23-24).  This partition can be seen in figure
\ref{sfig:part_fig:p_i}.  We similarly partition the shared subregion to
generate {\tt MAX\_PIECES} to correspond to the nodes that will be updated
by a each parallel computation.  This partitioning is also disjoint since
each node will only ever by updated by one parallel task and can be
seen in figure \ref{sfig:part_fig:s_i}.  Finally, we partition
the shared nodes again into the set of ghost nodes that we may want to make
multiple copies of since they will be read by multiple tasks.  Since
ghost nodes may be read by multiple parallel tasks, we may have to color
them multiple times and therefore the ghost partition is aliased
as can be seen in figure \ref{sfig:part_fig:g_i}.

Figure \ref{sfig:part_fig:tree} shows the final shape of the regions
and partitions.  The $*$ symbol indicates that a partition is 
disjoint.

@@

With the partitioning operations completed, the various subregions are
recorded in an array of {\tt CircuitPiece} structures.  Lines 32-33 fill in multiple
fields of a {\tt CircuitPiece} at once.  The need for such a statement will be described in detail in 
section \ref{subsec:datatypes}.

@@

Lines 35-39 form the bulk of the actual simulation, performing three passes
over the circuit on each iteration.  For each pass, a for loop is used to 
spawn a task for each piece of the circuit.  There are no explicit requests for
parallel execution, nor is there explicit synchronization required between the
passes.  Both the fact that the pieces can be run in parallel for a single 
pass and the the required inter-task dependencies are determined automatically
by the runtime based on the region access annotations on the task declarations.
We will describe how this process occurs in more detail in section 
\ref{subsec:cirdependence}.

The declarations for the three subtasks are shown on lines 42-51.  The 
{\tt calc\_new\_currents} task declares that it will read and write the region containing wires
for that piece, and will need to read both the region containing nodes in its piece and any
node that is adjacent to the piece (i.e. the ghost node region).  
The {\tt distribute\_charge} subtask turns things around, reading the piece's 
region of wires and updating all the nodes that those wires connect to.  However,
rather than requesting the ability to read and write the nodes (which would
require serialization of these tasks for correctness), the task declares that
it will use reorderable reduction operations and that the coherence requirement
can be reduced to atomicity rather than exclusive access. The final task is 
{\tt update\_voltages}, in which each piece is able to update
voltage information on the nodes contained in regions that it owns (i.e. private and shared)
in that piece based on the reductions that happened in the previous task.

@@

\subsection{Data Types and Constraints}
\label{subsec:datatypes}
There are two primitive data types for the simulation: nodes
and wires.  {\tt Node}s are linked in a circular list.
To reflect that pointers from one {\tt Node} to the next stay
in the same region we use the {\tt Type@region} syntax (e.g.
the {\tt next} field on line 1).
The program relies on different lists being in different
regions, and the {\tt Node} definition is parameterized on a region
{\tt rn} to reflect this constraint.  
The {\tt next} field of a $\tt Node \langle rn \rangle$ has
type $\tt Node \langle rn \rangle \mbox{\tt @} rn$, which implies that
all nodes in the list are in region {\tt rn}.

@@

In cases where we need to create compound data structures with types whose
region parameters are self-referential we declare a {\em region relationship}.
A region relationship is a structure with named fields, 
except that any region fields can be used in the type declarations
for other fields.  Lines 4-9 declare the region relationship {\tt Circuit}.
In {\tt Circuit} the region {\tt r\_all\_nodes} contains {\tt
Node}s that point only into region {\tt r\_all\_nodes}, guaranteeing that
all {\tt Node}s are in region {\tt r\_all\_nodes}.  (Note that regions are
declared with the single data type they can contain.)  Simiarly, the
type of the wire region's element not only guarantees that every wire
in the circuit is in the wire region, but also guarantees that every
node referred to by any wire is in the circuit's node region.

Since a region relationship can have
fields with types that refer its own fields' values, it is often not possible
to assign fields one at a time without violating the type checking rules.
The simultaneous assignment operator asks the type checker only to make sure
that the fields in the structure would have the right (self-referential) types
after all the fields have been changed.  Lines 32 and 33 illustrate a simultaneous
assignment to update a {\tt CircuitPiece} region relationship.

The last type definition is for the pieces into which the circuit is partitioned.
The original node and wire regions are specified as parameters, and the
four kinds of subregions that are created are made fields of the
{\tt CircuitPiece} (lines 10-17).  This declaration also illustrates {\em subregion constraints}
placed on the three region fields stating the region fields must be subregions of the 
node and wire regions provided as parameters (line 16).  Since the simulation code will need to be able to
iterate over the nodes owned by a {\tt CircuitPiece}, the parameterization of the {\tt rn\_pvt} (private nodes),
{\tt rn\_shr} (shared nodes), and {\tt first\_node} fields constrains the linked list of
nodes to reside entirely in those two regions.  The linked list of wires is
similarly constrained, and the wire's node pointers are constrained to fall
into one of the three subregions defined in the {\tt CircuitPiece}.  Finally,
the use of {\tt rn} for {\tt rn\_ghost}'s parameter and the second parameter
in the wire type places no constraints on which nodes a ghost node might point
to.

>>>>

The circuit simulation takes a collection of
of wires and nodes where wires meet.  
The simulation repeatedly updates
currents, distributes charges, and updates voltages for as many
time steps as the simulation demands.  We first
describe the partitioning of regions in Section~\ref{subsec:partitioning}
and then describe the data type declarations 
and the constraints they enforce 
in Section~\ref{subsec:datatypes}.

@@

Lines 19-27 are responsible for partitioning the circuit into {\tt MAX\_PIECES}
pieces that can be worked on in parallel.  To partition a region, we
provide a {\em coloring}, which is a relation between the elements
of a region and a set of colors.  A {\em partition} is an object which given
a coloring and a region $r$ contains for each color $c$ a {\em subregion} of $r$
containing all the elements of $r$ colored $c$.

Wires are simply partitioned in {\tt MAX\_PIECES} subregions (line 20).  
The nodes are first partitioned into 
{\em private} and {\em shared} nodes (line 21).  Private nodes
are touched by a single task in a phase, while shared nodes may
be referenced by multiple tasks in a phase.  The thick lines in figure \ref{sfig:part_fig:pvs}
show how the nodes in a small graph might be partitioned into three subsets.  Gray
nodes are shared and black nodes are private.  Note that
this partition is disjoint since each node has one color.
%Our current implementation uses small
%integers or booleans for the names of colors, but any enumerable set would be
%fine.  The coloring relation need not be total - it can leave out some elements entirely.  And although it is not required to be injective (it can map a single
%element to multiple colors), the fact that injective mappings will always
%result in disjoint subregions makes them strongly encouraged (when possible).
%Our example uses four colorings to create circuit partitions.

We next further partition the private and shared subregions into {\tt
MAX\_PIECES} disjoint subregions each (lines 23-24); one subregion of the private and shared node partitions
is shown in Figures~\ref{sfig:part_fig:p_i} and \ref{sfig:part_fig:s_i} respectively.  
We partition the shared nodes again
into the sets of ghost nodes for each task.  Since ghost nodes may be read by multiple tasks,
nodes may be colored multiple times and the ghost regions are not disjoint; one of the ghost subregions is shown in Figure~\ref{sfig:part_fig:g_i}.
Figure~\ref{sfig:part_fig:tree} shows the final hierarchy of node partitions.  The $*$ symbol indicates a partition is 
disjoint.

@@

%With the partitioning operations completed, the various subregions are
%recorded in an array of {\tt CircuitPiece} structures.  Lines 32-33 fill in multiple
%fields of a {\tt CircuitPiece} at once.  The need for such a statement will be described in detail in 
%section \ref{subsec:datatypes}.

@@

Lines 35-39 form the bulk of the actual simulation, performing three passes
over the circuit in each iteration.  Each pass loops over an array of pieces (constructed on lines 30-33 from the partitions),
spawning a task for each piece.   There are no explicit requests for
parallel execution, nor is there synchronization required between the
passes.  Both the fact that the tasks can be run in parallel within a pass
and the required inter-pass dependencies are determined automatically
by the Legion runtime based on the region access annotations on the task declarations
(see Section~\ref{subsec:cirdependence}).

%The declarations for the three subtasks are shown on lines 42-51.  
The  {\tt calc\_new\_currents} task reads and writes the wires subregion, 
the nodes subregion, and the ghost node region for its piece.
The {\tt distribute\_charge} subtask turns things around, reading the piece's 
wires subregion and updating all nodes those wires touch.  However,
rather than requesting the ability to read and write the nodes (which would
require serialization of these tasks for correctness), the task declares that
it will use reorderable {\em reduction} operations and that the coherence requirement
is {\em atomic} rather than exclusive access. The final task is 
{\tt update\_voltages}, which writes the shared and private nodes for a piece
and reads the results of the previous task's reductions.

@@

\subsection{Data Types and Constraints}
\label{subsec:datatypes}
The two basic data types are {\tt Node}s and {\tt Wire}s.
{\tt Node}s are linked in circular lists.
The simulation relies on different lists being in different
regions, so the {\tt Node} definition on line 1 is parameterized by a region
{\tt rn}.  The {\tt next} field of a $\tt Node \langle rn \rangle$ has
type $\tt Node \langle rn \rangle \mbox{\tt @} rn$, which means the field points
to an object in region {\tt rn} of type $\tt Node \langle rn \rangle$, 
implying that all nodes of the list are in region {\tt rn}.

@@

In cases where we need to create compound data structures with types whose
region parameters are self-referential we declare a {\em region relationship}.
A region relationship is a structure with named fields, 
except that any region fields can be used in the type declarations
for other fields.  Lines 4-9 declare the region relationship {\tt Circuit}.
In {\tt Circuit} the region {\tt r\_all\_nodes} contains {\tt
Node}s that point only into region {\tt r\_all\_nodes}, guaranteeing that
all {\tt Node}s are in region {\tt r\_all\_nodes}.  (Note that regions are
declared with the single data type they can contain.)  Similarly, the
type of the wire region's element not only guarantees that every wire
in the circuit is in the wire region, but also guarantees that every
node referred to by any wire is in the circuit's node region.

Since a region relationship can have
fields with types that refer its own fields' regions, it is often not possible
to assign fields one at a time without violating the type checking rules.
The simultaneous assignment operator asks the type checker only to ensure
that fields have the correct types after all the fields are updated.  
Lines 32 and 33 illustrate a simultaneous
assignment to update a {\tt CircuitPiece} region relationship.

The last type definition is for the pieces into which the circuit is partitioned.
The original node and wire regions are specified as parameters, and the
four kinds of subregions that are created are made fields of the
{\tt CircuitPiece} (lines 10-17).  This declaration also illustrates {\em subregion constraints}
placed on the three region fields stating the region fields must be subregions of the 
node and wire regions provided as parameters (line 16).  Since the simulation 
iterates over the nodes owned by a {\tt CircuitPiece}, the parameterization of the {\tt rn\_pvt} (private nodes),
{\tt rn\_shr} (shared nodes), and {\tt first\_node} fields constrains the linked list of
nodes to reside entirely in those two regions.  The linked list of wires is
similarly constrained, and the wire's node pointers are constrained to fall
into one of the three subregions.
% defined in the {\tt CircuitPiece}.  
Finally,
the use of {\tt rn} for {\tt rn\_ghost}'s parameter and the second parameter
in the wire type places no constraints on which nodes a ghost node might point
to.
